---
title: "mymain"
author: "Ethan Cook"
date: "2023-12-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Libraries
```{r}
library(dplyr)
library(ggplot2)
#library(recommenderlab)
#library(DT)
library(data.table)
library(reshape2)
library(coop)
```

# Load Data
## Ratings Data (ratings)
```{r}
myurl = "https://liangfgithub.github.io/MovieData/"

# use colClasses = 'NULL' to skip columns
ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
```

## Movies Data (movies)
```{r}
movies = readLines(paste0(myurl, 'movies.dat?raw=true'))
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)

# convert accented characters
movies$Title[73]
movies$Title = iconv(movies$Title, "latin1", "UTF-8")
movies$Title[73]

# extract year
movies$Year = as.numeric(unlist(
  lapply(movies$Title, function(x) substr(x, nchar(x)-4, nchar(x)-1))))
```

## User Data (users)
```{r}
users = read.csv(paste0(myurl, 'users.dat?raw=true'),
                 sep = ':', header = FALSE)
users = users[, -c(2,4,6,8)] # skip columns
colnames(users) = c('UserID', 'Gender', 'Age', 'Occupation', 'Zip-code')
```

## Ratings Matrix
```{r}
ratings_mat = read.csv("ratings_matrix.csv")
```

# System I: 

We propose a recommendation system that, given a genre, returns the the top ten highly-rated movies in that genre, given that there are more than 100 ratings for each movie that we are recommending. If there are less than 10 movies in that genre with more than 100 ratings, we append to that list with the top rated movies in general with no ratings minimum until the list is size 10.

```{r}
# Based on selected genre, chooses 10 popular recommendations
system_1_solver = function(genre)
{
  print(genre)
  
  ratings_df = ratings
  movies_df = movies
  
  # get movieID's of movies in that genre
  ids_in_genre = movies_df[grepl(genre, movies_df$Genres), "MovieID"]
  
  # If we have recommended for this genre already, load that saved data
  # instead of re-computing
  if (sum(saved_genre_recoms[[genre]]$MovieID) > 0)
  {
    ret = saved_genre_recoms[[genre]]
    print("reusing system1 genre recoms!")
  }
  else
  {
    ret = group_by(ratings_df, MovieID)
    ret = summarize(ret, ratings_per_movie = n(), ave_ratings = round(mean(Rating), dig=3))
    ret = inner_join(ret, movies_df, by = 'MovieID')
    ret = filter(ret, ratings_per_movie > 100)
    ret = filter(ret, MovieID %in% ids_in_genre)
    if (nrow(ret) < 10) {
      generic_recommendations = arrange(select(ratings_df, 'Rating', 'MovieID'), desc(Rating))[1:10,]
      ret = rbind(ret, generic_recommendations)
    }
    ret = top_n(ret, 10, ave_ratings)
    ret = select(ret, 'ave_ratings', 'MovieID')
    ret = arrange(ret, desc(-ave_ratings))  
    
    saved_genre_recoms[[genre]] <<- ret
  }
  ret
}
```

## STEP 1: Center Ratings Matrix 
```{r}
row_means = rowMeans(ratings_mat, na.rm=TRUE)
ratings_mat_centered = ratings_mat - row_means
```
## STEP 2: Cosine Similarity
```{r}
S = coop::cosine(as.matrix(ratings_mat_centered), use = "pairwise.complete.obs")
S = 0.5 + S/2
for (i in 1:nrow(S)) {
  S[i, i] = NA
  for (j in i:nrow(S)) {
    if ( length(intersect(which(!is.na(ratings_mat[,i])), which(!is.na(ratings_mat[,j])))) <= 2) {
      S[i, j] = NA
      S[j, i] = NA
    }
  }
}
write.csv(S, file="S_post_step2.csv")
```

## STEP 3: Top 30 only
```{r}
S_post_step2_ = read.csv("S_post_step2.csv")
S = S_post_step2_[,-1]
rownames(S) <- S_post_step2_[,1]

S[c("m1", "m10", "m100", "m1510", "m260", "m3212"), c("m1", "m10", "m100", "m1510", "m260", "m3212")]

for (r in 1:nrow(S)) {
  keep_idxs = order(as.matrix(S[r,]), decreasing=TRUE)[1:30]
  S[r,][-keep_idxs] = NA
}
write.csv(S, file="S_post_step3.csv")
```

!!!!
RUN CODE FROM HERE
!!!!

## STEP 4: myIBCF

```{r}
myIBCF = function(w) {
  recom_rankings = apply(S, 1, function(x){sum(x * w, na.rm=TRUE) / sum( (x * !is.na(w)), na.rm=TRUE)} )
  for (i in 1:length(recom_rankings))
  {
    if (!is.na(w[i]))
    {
      recom_rankings[i] = NA
    }
  }

  recom_rankings
}
```

```{r}
ratings_mat_with_rownames = read.csv("ratings_matrix.csv", row.names = 1)
```

## Test running IBCF
```{r}
S_post_step3_ = read.csv("S_post_step3.csv")
S = S_post_step3_[,-1]
rownames(S) <- S_post_step3_[,1]

start_time = Sys.time()
print("start time:")
print(start_time)

returned_rankings = myIBCF(ratings_mat_with_rownames["u1181",]) # dont bring to server.R

print("algorithm time:")
print(Sys.time() - start_time)

non_na_ranks = returned_rankings[!is.na(returned_rankings)]
head(non_na_ranks[order(-unlist(non_na_ranks))], n=20)

```

