---
title: "mymain"
author: "Ethan Cook"
date: "2023-12-02"
output: html_document
---

# Project 4
## Nishanth Alladi (nalladi2), Ethan Cook (elcook3), Davy Ji (davyji2)
Each member contributed to several pair programming sessions where the entirety of the report was written. Each member contributed to both theorizing solutions and in implementing code.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Libraries
```{r message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(data.table)
library(reshape2)
library(coop)
```

## Load Data
### Ratings Data (ratings)
```{r}
myurl = "https://liangfgithub.github.io/MovieData/"

# use colClasses = 'NULL' to skip columns
ratings = read.csv(paste0(myurl, 'ratings.dat?raw=true'), 
                   sep = ':',
                   colClasses = c('integer', 'NULL'), 
                   header = FALSE)
colnames(ratings) = c('UserID', 'MovieID', 'Rating', 'Timestamp')
```

### Movies Data (movies)
```{r}
movies = readLines(paste0(myurl, 'movies.dat?raw=true'))
movies = strsplit(movies, split = "::", fixed = TRUE, useBytes = TRUE)
movies = matrix(unlist(movies), ncol = 3, byrow = TRUE)
movies = data.frame(movies, stringsAsFactors = FALSE)
colnames(movies) = c('MovieID', 'Title', 'Genres')
movies$MovieID = as.integer(movies$MovieID)

# convert accented characters
movies$Title = iconv(movies$Title, "latin1", "UTF-8")

# extract year
movies$Year = as.numeric(unlist(
  lapply(movies$Title, function(x) substr(x, nchar(x)-4, nchar(x)-1))))
```

### User Data (users)
```{r}
users = read.csv(paste0(myurl, 'users.dat?raw=true'),
                 sep = ':', header = FALSE)
users = users[, -c(2,4,6,8)] # skip columns
colnames(users) = c('UserID', 'Gender', 'Age', 'Occupation', 'Zip-code')
```

### Ratings Matrix
```{r}
ratings_mat = read.csv("ratings_matrix.csv")
```

# System I: 

We propose a recommendation system that, given a genre, returns the the top ten highly-rated movies in that genre, given that there are more than 100 ratings for each movie that we are recommending. If there are less than 10 movies in that genre with more than 100 ratings, we append to that list with the top rated movies in general with no ratings minimum until the list is size 10. Below we have provided a code implementation of the proposal as it runs on our webpage.

We also save the top 10 recommendations for any given genre once they've already been calculated, so if the same genre is selected again, no computation is needed.

```{r}
# Based on selected genre, chooses 10 popular recommendations
system_1_solver = function(genre)
{
  print(genre)
  
  ratings_df = ratings
  movies_df = movies
  
  # get movieID's of movies in that genre
  ids_in_genre = movies_df[grepl(genre, movies_df$Genres), "MovieID"]
  
  # If we have recommended for this genre already, load that saved data
  # instead of re-computing
  if (sum(saved_genre_recoms[[genre]]$MovieID) > 0)
  {
    ret = saved_genre_recoms[[genre]]
    print("reusing system1 genre recoms!")
  }
  else
  {
    ret = group_by(ratings_df, MovieID)
    ret = summarize(ret, ratings_per_movie = n(), ave_ratings = round(mean(Rating), dig=3))
    ret = inner_join(ret, movies_df, by = 'MovieID')
    ret = filter(ret, ratings_per_movie > 100)
    ret = filter(ret, MovieID %in% ids_in_genre)
    if (nrow(ret) < 10) {
      generic_recommendations = arrange(select(ratings_df, 'Rating', 'MovieID'), desc(Rating))[1:10,]
      ret = rbind(ret, generic_recommendations)
    }
    ret = top_n(ret, 10, ave_ratings)
    ret = select(ret, 'ave_ratings', 'MovieID')
    ret = arrange(ret, desc(-ave_ratings))  
    
    saved_genre_recoms[[genre]] <<- ret
  }
  ret
}
```

## STEP 1: Center Ratings Matrix 
```{r}
row_means = rowMeans(ratings_mat, na.rm=TRUE)
ratings_mat_centered = ratings_mat - row_means
```

## STEP 2: Cosine Similarity
```{r}
S = coop::cosine(as.matrix(ratings_mat_centered), use = "pairwise.complete.obs")
S = 0.5 + S/2
for (i in 1:nrow(S)) {
  S[i, i] = NA
  for (j in i:nrow(S)) {
    if ( length(intersect(which(!is.na(ratings_mat[,i])), which(!is.na(ratings_mat[,j])))) <= 2) {
      S[i, j] = NA
      S[j, i] = NA
    }
  }
}
write.csv(S, file="S_post_step2.csv")
```

## STEP 3: Top 30 only
```{r}
S_post_step2_ = read.csv("S_post_step2.csv")
S = S_post_step2_[,-1]
rownames(S) <- S_post_step2_[,1]

S[c("m1", "m10", "m100", "m1510", "m260", "m3212"), c("m1", "m10", "m100", "m1510", "m260", "m3212")]

for (r in 1:nrow(S)) {
  keep_idxs = order(as.matrix(S[r,]), decreasing=TRUE)[1:30]
  S[r,][-keep_idxs] = NA
}
write.csv(S, file="S_post_step3.csv")
```

## STEP 4: myIBCF
```{r}
myIBCF = function(w) {
  recom_rankings = apply(S, 1, function(x){sum(x * w, na.rm=TRUE) / sum( (x * !is.na(w)), na.rm=TRUE)} )
  for (i in 1:length(recom_rankings))
  {
    if (!is.na(w[i]))
    {
      recom_rankings[i] = NA
    }
  }
  recom_rankings
}
```

### load data for testing IBCF
```{r}
ratings_mat_with_rownames = read.csv("ratings_matrix.csv", row.names = 1)

S_post_step3_ = read.csv("S_post_step3.csv")
S = S_post_step3_[,-1]
rownames(S) <- S_post_step3_[,1]
```

### These functions run on the server
If less than 10 recommendations are returned from system 2, we choose movies with high average ratings from the movies data that are not related to the users ratings, and fill in the missing recommendations.
```{r}
process_user_rating_output = function(user_ratings)
{
  # Pull a single sample user from ratings matrix and set all values to NA
  ratings_mat_sample = read.csv("ratings_matrix.csv", nrows = 1)
  ratings_mat_sample[!is.na(ratings_mat_sample)] = NA

  # need to remove strange NA movie id
  temp = user_ratings[-1,]

  movie_ids = temp$MovieID
  for (id in 1:length(movie_ids))
  {
    text_id = paste("m", as.character(movie_ids[id]), sep="")
    ratings_mat_sample[text_id] = temp$Rating[id]
  }
  # convert to numeric vector to save processing time by like 30,000%
  as.numeric(ratings_mat_sample)
}

print_top_10 = function(returned_rankings)
{
  non_na_ranks = returned_rankings[!is.na(returned_rankings)]
  
  # If less than 10 recommendations, put some other movies in
  if (length(non_na_ranks) < 10) {
    nrank = 10 - length(non_na_ranks)
    generic_recommendations = arrange(select(ratings_df, 'Rating', 'MovieID'), desc(Rating))[1:nrank,]
    vector_generics = rep(1, nrank)#select(generic_recommendations, 'Rating')[['Rating']]
    vec_names = select(generic_recommendations, 'MovieID')[['MovieID']]
    for (id in 1:length(vec_names))
    {
      vec_names[id] = paste("m", as.character(vec_names[id]), sep="")
    }
    names(vector_generics) = vec_names
    full_ranks = c(non_na_ranks, vector_generics)
    print("less than 10 recoms, injecting...")
  }
  else
  {
    full_ranks = non_na_ranks
  }
  
  # format results as a dataframe with MovieIDs and Scores
  return_df = data.frame(
    ids = names(head(full_ranks[order(-unlist(full_ranks))], n=10)),
    scores = head(full_ranks[order(-unlist(full_ranks))], n=10)
  )
  print(return_df)
}
```

### Test user 1181
```{r}
returned_rankings = myIBCF(as.numeric(ratings_mat_with_rownames["u1181",])) # dont bring to server.R
print_top_10(returned_rankings)
```

### Test user 1351
```{r}
returned_rankings = myIBCF(as.numeric(ratings_mat_with_rownames["u1351",])) # dont bring to server.R
print_top_10(returned_rankings)
```

### Create and test hypothetical user
```{r}
ratings_mat_sample = read.csv("ratings_matrix.csv", nrows = 1)
ratings_mat_sample[!is.na(ratings_mat_sample)] = NA
ratings_mat_sample["m1613"] = 5
ratings_mat_sample["m1755"] = 4
returned_rankings = myIBCF(as.numeric(ratings_mat_sample)) # dont bring to server.R
print_top_10(returned_rankings)
```



